Results publications are inadequately linked to trial registrations: An automated pipeline and evaluation of German university medical centers

### 2021-08-02 docs/manuscript_papaja.Rmd
Update manuscript

abstract: |
  **Objective** To evaluate, using a novel automated approach, the links between registrations and results publications across a sample of clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS). The World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linking to make both clinical trial publications and registrations findable and thus increase the value of clinical research.

@@

  **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials comprehensively link their registrations and publications, meaning a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata. In contrast, `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` trials had no links. The most common link was TRN reported in the full-text, which occurred for `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_trn_ft")` trials. ClinicalTrials.gov trials were overall better linked than DRKS trials, and this difference appears to be driven by PubMed and registry infrastructure, rather than by trialists. Of trials reporting a TRN in the abstract, ClinicalTrials.gov trials were more likely than DRKS trials to have the TRN captured in the PubMed metadata. Additionally, most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_ctgov_link_reg_trn_si_derived")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.

@@

keywords          : "clinical trials, registration, reporting, meta-research"
wordcount         : "3714"

@@

header-includes   :
  - \usepackage{orcidlink}
---

@@

library(ggplot2)
library(DiagrammeR)
library(gtsummary)
library(gt)
library(papaja)

@@

# Prepare references
r_refs("r-references.bib")
r_citations <- 
  cite_r(file = "r-references.bib", footnote = TRUE) # pkgs, withhold #r_citations$r and r_citations$pkgs

@@

```{r abstract-vars, cache = TRUE}
abs_n_trials_link_full <- report_pct_num_denom_trials(n_trials_link_full)
abs_n_trials_link_none <- report_n_pct_trials(n_trials_link_none)
abs_n_trials_link_trn_ft <- report_n_pct_trials(n_trials_link_trn_ft)

@@

Clinical trial results risk being left out of evidence synthesis and health policy decision making if publications cannot be found. Incomplete evidence threatens healthcare and the lives and welfare of human volunteers. Legal regulations as well as ethical guidelines call for trial results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp]. Linking trial registrations and results publications makes each more findable, reducing the risk of information loss and increasing clinical research value [@kubben2019] <!--# looked for pub on FAIR/finable clinical trials and not finding much beyond just data-->. Furthermore, such threaded evidence empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd].

@@

Trial evidence appears to be insufficiently linked in both registries and publications. Previous studies on trial registrations found links to results publications in as few as 13% to as many as 42% of trials [@bashir2017sr; @huser2013po]. TRN reporting in either the metadata, abstract, or full-text of trial publications has been found to be as low at 8% and as high as 97%, varying widely depending on how the trial population was defined [@bashir2017sr; @carlisle2020m; @aldurra2020b]. For example, some studies included trial publications limited to specific journals or fields, whereas others relied on PubMed "clinical trial" publication type queries.

In order to draw robust conclusions about registration-publication linking, the analysis population must be limited to registered clinical trials with published results. Previous projects faced limitations due to the unclear publication cohort and unknown true registration rates. PubMed queries currently do not suffice to identify the true populations of clinical trials or even RCTs, both excluding true positives (i.e., sensitivity) and including false positives (i.e., specificity) [@glanville2020jmla]. Therefore not all PubMed-indexed clinical trials or RCTs should be registered <!--# as they may not reflect true clinical trials or RCTs -->, and not all that should be are. Furthermore, not all reported TRNs correspond to bona fide registrations of the trial whose results are reported in the publication, i.e., TRNs may be reported for related trials in the background section. As such, conclusions drawn by these previous studies relying on PubMed-query based samples, about registration reporting rates must be qualified as the rate of reported trial registration numbers in PubMed-indexed (randomized) clinical trials with unknown registration rates. <!--# remove previous sentence? --> Similarly, registry-based cohorts have included trials that may not have an associated publication and thus should not be expected to have a publication link. Furthermore, these studies of publication links in registrations have focused on ClinicalTrials.gov [@huser2013po; @huser2012; @goodwin2018], and we are not aware of previous studies investigating publication links in DRKS.

@@

We used two cohorts of registered clinical trials and associated results previously developed by @wieschowski2019joce and @riedel2021 and referred to as the "IntoValue" dataset. In brief, the dataset consists of clinical trials registered on ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) and conducted at a German UMC completed between 2009 and 2017, and results publications found via manual searches [ADD DATASET REFERENCE].

@@

We queried the PubMed Entrez Programming Utilities (E-utilities) API<!--#on DAY MONTH YEEAR--> for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @winter]. From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata.<!--# , as well as MeSH terms, keywords, author details, and summary metadata -->

We used DOIs and PMIDs to search for full-text publications as PDFs, using a combination of automated and manual strategies, including contacting the corresponding author as a final step <!--# add that this was just for a few studies? -->[@ourresearch; @jahn]. We then used the GROBID machine learning library for technical and scientific publications (v0.6.1) via Python to parse the PDFs into machine-readable XMLs [@2008; @lopez2021]. In order to isolate the main body from the publication abstract for our analysis, we extracted the `<body>` sections of the papers.

Figure \@ref(fig:flow-methods) show the methods flow using automated techniques<!--# with the exception of some full-text PDFs -->.

```{r flow-methods, fig.cap = "Data collection sources and methods", out.width="50%"}
knitr::include_graphics(here("docs", "figures", "flow-methods.png"))
```

@@

After updating registry data, we reapplied the IntoValue exclusion criteria to exclude trials that were completed before 2009 or after 2017, were non-interventional, or did not meet the recruitment or study status inclusion criteria (see @wieschowski2019joce and @riedel2021 for further details on these criteria). We also excluded trials without a German UMC lead, as only one of the two IntoValue cohorts included trials not led by German UMCs. For duplicate trials (i.e., appearing in both IntoValue cohorts), we excluded the earlier record. We also excluded trials without results, and trials with results for which we could not find a PMID and full-text publication.

@@

As reference data is available in difference formats depending on the registry, we developed parallel approaches to extract publication identifiers for each. For ClinicalTrials.gov, we retrieved the reference type, citation, and PMID fields from AACT. We then used a regular expression (regex) to extract any DOIs from the citations; we did not search for PMIDs as these are captured in an AACT field. For DRKS, we scraped the reference type, citation, and URL, when available. We used regexes to extract any DOIs and PMIDs from the citation and URL. In the rare cases when conflicting DOIs or PMIDs were found in the citation vs. the URL, we manually reviewed to determine which was valid; if both were valid, we preferred the identifier provided in the citation. In addition, we determined whether a reference was manually or automatically indexed in the registry. DRKS allows only manually added references; for ClinicalTrials.gov, references of type "derived" were marked as automatically indexed. We then used DOIs and PMIDs to match the referenced publications to those in our trial dataset. We did not attempt to match publications without identifiers (i.e., publication title only).

@@

We developed <!--# and evaluated [or more methods details] -->regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (see Appendix \@ref(tab:tbl-trn-regexes)) <!--# or move to osf protocol or remove from this paper and make separate citation  --> and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text. To gauge the sensitivity and specificity of these regexes, we visually inspected all PubMed secondary identifier metadata<!--# for true positives and true negatives-->. For each source, we first extracted all unique TRN patterns within a source. Our regex allowed for minor formatting errors in the TRN patterns (such as additional punctuation). We then cleaned the TRN patterns to remove these formatting errors, and then deduplicated the corrected TRNs to exclude duplicates within a source uncovered through the cleaning process. We also coalesced across sources to deduce a list of unique TRNs by publication along with the source(s) in which each was reported.

The output of this TRN extraction, cleaning, and deduplication included TRNs beyond those of the known registrations. These additional TRNs could be cross-registrations of the same trial or provided as background or discussion. For this study, we were interested in whether known registrations were reported in the publication and so filtered for reported TRNs matching the known registrations. <!--# change this if considering cross-reg -->

@@

We generated descriptive statistics on trial and publication characteristics, overall and by registry. We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration.<!--#with each type of link and different combination of links.--> To explore change over time, the relationships between types of links, and differences between registries, we ran exploratory logistic regressions for each link type, with all other link types as well as registry and completion year as explanatory variables. In particular, since PubMed metadata may be generated from the publication abstract or full-text, we examined the relationship between TRN reporting in either the abstract or full-text, and TRN inclusion in the metadata. Additionally, since ClinicalTrials.gov registrations automatically reference publications with the TRN in the PubMed metadata (whereas DRKS does not), we examined the proportion of automatically vs. manually linked publications in ClinicalTrials.gov. To explore registry differences for only manually linked references, we excluded automated links and calculated the number and proportion of trials with the publication linked in the registration. <!--#add automatic links to regression for pub in registry? in which case would need to recode NA as FALSE. To explore whether automated links explains registry differences in publication links in the registry...-->

@@

Data collection, preparation, and analysis was performed in `r r_citations$r`<!--#aside from parsing PDFs to XML in Python-->.<!--#Additional packages and libraries are specified in the methods.--> All code and the final analysis data for this study are available at <https://github.com/maia-sh/reg-pub-link>. Raw data (with the exception of the full-text of publications) is available at <https://github.com/maia-sh/intovalue-data>. `r r_citations$pkgs`

@@

The IntoValue dataset includes all trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`). After applying our exclusion criteria, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial. Figure \@ref(fig:flow-trials) details the trial and publication screening. Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications by registry. <!--# QUESTION: what summary info about trials and publications? I also include a long version of demographic data in case we want to include any of these, ideally with collapsed categories, in Table @ref(tab:tbl-trials-full) -->

@@

```{r tbl-trials}
tbl_trials %>% 
  modify_caption("Characteristics of included trials with results publications")
  # gtsummary::as_flex_table() %>% 
  # flextable::set_caption(caption = "Characteristics of included trials with results publications")
```

@@

```{r tbl-link-by-reg}
tbl_link_by_reg %>% 
  modify_caption("Registration-publication links by registry")
  # gtsummary::as_flex_table() %>% 
  # flextable::set_caption(caption = "Registration-publication links by registry")
```

@@

```{r tbl-trials-regression}
tbl_trials_regression %>% 
  modify_caption("Crude and adjuusted odds ratios for factors associated with publication-registration links")
  # gtsummary::as_flex_table() %>% 
  # flextable::set_caption(caption = "Crude and adjuusted odds ratios for factors associated with publication-registration links")
```

Table \@ref(tab:tbl-trials-regression) shows the crude univariate and adjusted multivariate odds ratios (ORs) for each type of publication-registration link across all explanatory variables. Completion year did not have a strong relationship with linking practices, although trials completing more recently were more likely report the TRN in the abstract (`r inline_text(cor_abs, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(cor_ft, variable = "completion_year", pattern = "cOR {estimate} [{ci}]")`; `r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`). Figure \@ref(fig:plot-links-year-reg) shows the rate of registration-publication links over time for ClinicalTrials.gov and DRKS.

@@

Across the adjusted multivariate models,<!--#in the adjusted analysis,--> trials with one type of link were generally more likely to have another type of link. <!--#However, trials with a TRN in the full-text were significantly less likely to have the publication linked in the registration (`r inline_text(aor_reg, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")` ; `r inline_text(aor_ft, variable = "has_reg_pub_link", pattern = "aOR {estimate} [{ci}]")`).--> In particular, TRNs were more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")`). Futhermore, DRKS TRNs were less likely than ClinicalTrials.gov TRNs to appear in the PubMed metadata (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`). Similarly, for trials with a TRN in either abstract or full-text, ClinicalTrials.gov TRNs appeared at a higher rate than DRKS TRNs in the metadata (`r inline_text(tbl_si_abs_or_ft, "has_iv_trn_secondary_id", column = "stat_1", pattern = "{p}%")` vs. `r inline_text(tbl_si_abs_or_ft, "has_iv_trn_secondary_id", column = "stat_2", pattern = "{p}%")`). Table \@ref(tab:tbl-si-if-abs-ft) shows the number and proportion of trials with a TRN in the metadata given a TRN in the abstract, full-text, or either, both overall and by registry.

```{r tbl-si-if-abs-ft}
tbl_si_if_abs_ft
```

@@

Linking of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of this resource-intensive area of biomedical research. Poor linking poses a barrier to identifying trial publications via automated approaches and instead require researchers to perform intensive manual keyword-based searches to match publications to trials [@wieschowski2019joce; @chen2016b; @bashir2017sr]. This responsible practice comes at minimal costs to researchers, from seconds for pasting TRNs in papers, to minutes for adding a publication link to the registration.

@@

This study reveals shortcomings in the linking of German UMC-led trials registrations and results publications and highlights several promising avenues forward. In contrast with other responsible clinical research practices (such as data protection or trial registration itself), registration-publication linking is relatively straightforward and can be improved with action across stakeholders. Drawing from the study findings and additional context, Table  \@ref(fig:tbl-stakeholders) outlines recommendations for stakeholders across clinical research.

```{r tbl-stakeholders}
# https://www.anthonyschmidt.co/post/2020-06-03-making-apa-tables-with-gt/
# results = "asis"
# tbl_stakeholders %>%
#   as_latex()
knitr::include_graphics(here::here("docs", "figures", "tbl-stakeholders.png"))
```

@@

We would like to thank Kerstin Rubarth from the Charité Institut für Biometrie und Klinische Epidemiologie for her consultation on the statistical analysis.

@@

# ```{r echo = FALSE, results = 'asis', cache = FALSE}
# papaja::render_appendix('appendix.Rmd')
# ```

>>>>

abstract: |
  **Objective** To evaluate, using a novel automated approach, links between registration and publication across clinical trials led by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS). The World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linking to make clinical trial publications and registrations findable and increase the value of clinical research.

@@

  **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials have a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata. In contrast, `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` trials had no links. The most common link was TRN reported in the full-text (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_trn_ft")`). ClinicalTrials.gov trials were overall better linked than DRKS trials, and this difference appears to be driven by PubMed and registry infrastructure, rather than by trialists. Of trials reporting a TRN in the abstract, trials in ClinicalTrials.gov were more likely than trials in DRKS to have the TRN captured in the PubMed metadata. Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_ctgov_link_reg_trn_si_derived")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.

@@

keywords          : "clinical trials, registration, reporting, meta-research"
wordcount         : "3551"

@@

header-includes   :
  - \usepackage{orcidlink}

always_allow_html: true
---

@@

library(ggplot2)
library(DiagrammeR)
library(gtsummary)
library(gt)
library(papaja)
# library(knitr)
# library(kableExtra)

library(rentrez)
library(roadoi)
theme_gtsummary_compact()

@@

# Prepare references
r_refs("r-references.bib")
r_citation <- cite_r() # file = "r-references.bib", footnote = TRUE, pkgs, withhold #r_citations$r and r_citations$pkgs

@@

```{r abstract-vars, cache = TRUE}
abs_n_trials_link_full <- report_pct_num_denom_trials(n_trials_link_full)
abs_n_trials_link_none <- report_n_pct_trials(n_trials_link_none)
abs_n_trials_link_trn_ft <- "1131, 60% of trials"
  # Change to gtsummary once updated: https://stackoverflow.com/questions/68502977/gtsummaryinline-text-for-stat-0-with-pattern
  # inline_text(tbl_link_by_reg, has_iv_trn_ft_pdf, stat_0, pattern = "{n}, {p}% of trials")
  # report_n_pct_trials(n_trials_link_trn_ft)

@@

Clinical trial evidence informs clinical practice and health policy decision making through reviews and meta-analyses. The validity of health evidence synthesis is jeopardized if relevant trial data are left out to do inadequate or incomplete trial metadata, or inability to find trial results through automated means. Incomplete evidence threatens healthcare and the lives and welfare of human volunteers. Legal regulations as well as ethical guidelines call for trial results transparency [@worldmedicalassociation2013; @worldhealthorganization2017; @altman2014gfrhraum; @borysowski2020bjcp; @chalmers2013b]. Linking trial registrations and results publications makes each more findable, reducing the risk of information loss and increasing clinical research value [@kubben2019] <!--# looked for pub on FAIR/finable clinical trials and not finding much beyond just data-->. Furthermore, such threaded evidence empowers readers to cross-check sources, e.g., for potential outcome switching [@altman2014t], and can support systematic reviewers who rely on registrations to supply additional trial information for evidence synthesis [@miron2020sd].

@@

Trial evidence appears to be insufficiently linked in both registrations and publications. Previous studies on trial registrations found links to results publications in as few as 13% to as many as 60% of trials [@bashir2017sr; @huser2013po; @wetering2012po]. TRN reporting in either the metadata, abstract, or full-text of trial publications has been found to be as low at 8% and as high as 97%, varying widely depending on how the trial population was defined [@bashir2017sr; @carlisle2020m; @aldurra2020b; @huser2013jamia; @wetering2012po]. The sampling strategies used in these previous studies limit our ability to draw conclusions about registration-publication linking. Publication-based cohorts relied on inaccurate PubMed clinical trial filters [@glanville2020jmla] and included non-trial-result publications, and trials with unknown registration rates (as not all trials are registered, [@trinquart2018bm; @gopal2018t; @denneny2019bo]); registration-based cohorts included included trials that may not have an associated publication and thus should not be expected to have a publication link [@chen2016b; @dwan2013po; @ross2012b]. Furthermore, these studies of publication links in registrations have primarily focused on ClinicalTrials.gov, and we are not aware of previous studies investigating publication links in DRKS. In this study, we limit our sample to registered clinical trials with published results which can be expected to link the registration and publication, and which allows us to draw conclusions about the prevalence of this responsible research practices.

@@

We used two cohorts of registered clinical trials and associated results previously developed by @wieschowski2019joce and @riedel2021 and referred to as the "IntoValue" dataset [@riedel2021a]. In brief, the dataset consists of clinical trials registered on ClinicalTrials.gov or the German Clinical Trials Registry (DRKS) and conducted at a German UMC completed between 2009 and 2017, and results publications found via manual searches.

@@

We queried the PubMed Entrez Programming Utilities (E-utilities) API<!--#on DAY MONTH YEAR--> for all trial results PMIDs, and excluded PMIDs that did not resolve [@usnationallibraryofmedicinetigtand; @R-rentrez]. From the PubMed XML, we extracted bibliometric information, including the publication abstract and secondary identifier (or databank) metadata.<!--# , as well as MeSH terms, keywords, author details, and summary metadata -->

We used DOIs and PMIDs to search for full-text publications as PDFs, using a combination of automated and manual strategies, including contacting the corresponding author as a final step <!--# add that this was just for a few studies? -->[@ourresearch; @R-roadoi]. We then used the GROBID machine learning library for technical and scientific publications (v0.6.1) via Python to parse the PDFs into machine-readable XMLs [@2008; @lopez2021]. In order to isolate the main body from the publication abstract for our analysis, we extracted the `<body>` sections of the papers.

@@

After updating registry data, we reapplied the IntoValue exclusion criteria (completion date before 2009 or after 2017, non-interventional, incomplete based on study status, non-German-UMC lead; see @riedel2021 for further details on these criteria) and limited our sample to trials with results with a PMID and full-text publication.

@@

As reference data is available in different formats depending on the registry, we developed parallel approaches to extract publication identifiers for each. For ClinicalTrials.gov, we retrieved the reference type, citation, and PMID fields from AACT. We then used a regular expression (regex) to extract any DOIs from the citations; we did not search for PMIDs as these are captured in an AACT field. For DRKS, we scraped the reference type, citation, and URL, when available. We used regexes to extract any DOIs and PMIDs from the citation and URL. In the rare cases when conflicting DOIs or PMIDs were found in the citation vs. the URL, we manually reviewed to determine which was valid; if both were valid, we preferred the identifier provided in the citation. In addition, we determined whether a reference was manually or automatically indexed in the registry. DRKS allows only manually added references; for ClinicalTrials.gov, references of type "derived" were marked as automatically indexed. We then used DOIs and PMIDs to match the referenced publications to those in our trial dataset. We did not attempt to match publications without identifiers (i.e., publication title only).

@@

We developed <!--# and evaluated [or more methods details] -->regexes for the TRN patterns for all PubMed-indexed and ICTRP-network registries (available at <https://github.com/maia-sh/ctregistries/blob/master/inst/extdata/registries.csv>) and used these to detect and classify TRNs in the PubMed secondary identifier metadata, abstract, and full-text. To gauge the sensitivity and specificity of these regexes, we visually inspected all PubMed secondary identifier metadata<!--# for true positives and true negatives-->. For each source, we first extracted all unique TRN patterns within a source. Our regex allowed for minor formatting errors in the TRN patterns (such as additional punctuation). We then cleaned the TRN patterns to remove these formatting errors, and then deduplicated the corrected TRNs to exclude duplicates within a source uncovered through the cleaning process. We also merged our sources to produce a list of unique TRNs by publication along with the source(s) in which each was reported.

The output of TRN extraction, cleaning, and deduplication included TRNs beyond those of the known registrations. These additional TRNs could be cross-registrations of the same trial or provided as background or discussion. For this study, we were interested in whether known registrations were reported in the publication and so filtered for reported TRNs matching the known registrations. <!--# change this if considering cross-reg -->

@@

We generated descriptive statistics on trial and publication characteristics, overall and by registry. We calculated the number and proportion of trials linked via the publication (full-text, abstract, metadata) and the registration.<!--#with each type of link and different combination of links.--> To explore change over time, the relationships between types of links, and differences between registries, we ran logistic regressions for each link type, with all other link types as well as registry and completion year as explanatory variables. Regressions were performed as exploratory analyses, and all variables were included in each regression model.<!--#no model selection or fitting was conducted; the explanatory variables were tested a priori to verify there was no violation of the assumption of the linearity of the logit TODO, ask Murph--> In particular, since PubMed metadata may be generated from the publication abstract or full-text, we examined the relationship between TRN reporting in either the abstract or full-text, and TRN inclusion in the metadata. Additionally, since ClinicalTrials.gov registrations automatically reference publications with the TRN in the PubMed metadata (whereas DRKS does not), we examined the proportion of automatically vs. manually linked publications in ClinicalTrials.gov. To explore registry differences for only manually linked references, we excluded automated links and calculated the number and proportion of trials with the publication linked in the registration. <!--#add automatic links to regression for pub in registry? in which case would need to recode NA as FALSE. To explore whether automated links explains registry differences in publication links in the registry...-->

@@

Data collection, preparation, and analysis was performed in `r r_citation`<!--#aside from parsing PDFs to XML in Python-->.<!--#Additional packages and libraries are specified in the methods.--> All code, the final analysis data, and a list of all packages used for this study are available at <https://github.com/maia-sh/reg-pub-link>. Raw data (with the exception of the full-text of publications) is available at <https://github.com/maia-sh/intovalue-data>.

@@

The IntoValue dataset includes all trials led by a German UMC that were registered on ClinicalTrials.gov or DRKS and completed between 2009 and 2017 (n = `r n_trials_iv`). After applying our exclusion criteria, our sample included `r n_trials_ft_pdf` trials with `r n_pubs_unique` unique results publications indexed in PubMed and available as full-text, as some publications include results from more than one trial. Figure \@ref(fig:flow-trials) details the trial and publication screening. Table \@ref(tab:tbl-trials) shows summary descriptive details of the trials with results publications by registry.

@@

<!-- (ref:tbl-trials-caption) **Characteristics of German UMC-led trials with published results** A trial was considered randomized if allocation included randomization. A trial was considered prospectively registered if registered in the same or previous months to start date. Summary results were taken from a structured data field in ClinicalTrials.gov, and determined based on manual inspection for terms such as *Ergebnisbericht* or *Abschlussbericht* in DRKS. *Top* journals refer to the journals with the greatest number of trial publications in our sample. -->

```{r tbl-trials, results = "asis"}
# tbl <- read_rds(here("docs", "figures", "tbl-trials.rds"))
# tbl %>% 
#   as_kable_extra(booktabs = TRUE)
tbl_trials %>%

  as_kable_extra()
  # as_tibble() %>%
  # apa_table(
  #   caption = "This is a table about trials",
  #   format.args = list(na_string = ""),
  #   format = "pandoc"
  # )
# modify_caption("Characteristics of included trials with results publications")
  # gtsummary::as_flex_table() %>% 
  # flextable::set_caption(caption = "Characteristics of included trials with results publications")
```

@@

```{r tbl-link-by-reg}
tbl_link_by_reg %>% 
  modify_caption("Registration-publication links by registry") #%>% 
  # gtsummary::as_flex_table() #%>%
  # flextable::set_caption(caption = "Registration-publication links by registry")
```

@@

```{r tbl-trials-regression}
tbl_trials_regression %>% 
  modify_caption("Crude and adjusted odds ratios for factors associated with publication-registration links")
  # gtsummary::as_flex_table() %>% 
  # flextable::set_caption(caption = "Crude and adjuusted odds ratios for factors associated with publication-registration links")
```

Table \@ref(tab:tbl-trials-regression) shows the crude univariate and adjusted multivariate odds ratios (ORs) for each type of publication-registration link across all explanatory variables. Completion year did not have a strong relationship with linking practices, although trials completing more recently were more likely report the TRN in the abstract (`r inline_text(aor_abs, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`) and full-text (`r inline_text(aor_ft, variable = "completion_year", pattern = "aOR {estimate} [{ci}]")`). Figure \@ref(fig:plot-links-year-reg) shows the rate of registration-publication links over time for ClinicalTrials.gov and DRKS.

@@

Across the adjusted multivariate models,<!--#in the adjusted analysis,--> trials with one type of link were generally more likely to have another type of link. <!--#However, trials with a TRN in the full-text were significantly less likely to have the publication linked in the registration (`r inline_text(aor_reg, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")` ; `r inline_text(aor_ft, variable = "has_reg_pub_link", pattern = "aOR {estimate} [{ci}]")`).--> In particular, TRNs were more likely to appear in the PubMed metadata if included in the abstract (`r inline_text(aor_si, variable = "has_iv_trn_abstract", pattern = "aOR {estimate} [{ci}]")`), but not more likely if included in the full-text (`r inline_text(aor_si, variable = "has_iv_trn_ft_pdf", pattern = "aOR {estimate} [{ci}]")`). Futhermore, DRKS TRNs were less likely than ClinicalTrials.gov TRNs to appear in the PubMed metadata (`r inline_text(aor_si, variable = "registry", level = "DRKS", pattern = "aOR {estimate} [{ci}]")`). Similarly, for trials with a TRN in either abstract or full-text, ClinicalTrials.gov TRNs appeared at a higher rate than DRKS TRNs in the metadata (`r inline_text(tbl_si_abs_or_ft, "has_iv_trn_secondary_id", column = "stat_1", pattern = "{p}%")` vs. `r inline_text(tbl_si_abs_or_ft, "has_iv_trn_secondary_id", column = "stat_2", pattern = "{p}%")`). Appendix \@ref(si-if-abs-ft) shows the number and proportion of trials with a TRN in the metadata given a TRN in the abstract, full-text, or either, both overall and by registry.

@@

Linking of trial registrations and results publications plays an important role in research transparency and can substantially increase the value of this resource-intensive area of biomedical research. Poor linking poses a barrier to identifying trial publications via automated approaches and instead require researchers to perform intensive manual keyword-based searches to match publications to trials [@wieschowski2019joce; @chen2016b; @bashir2017sr]. This responsible research practice comes at minimal costs to researchers, from seconds for pasting TRNs in papers, to minutes for adding a publication link to the registration.

@@

This study reveals shortcomings in the linking of German UMC-led trials registrations and results publications and highlights several promising avenues forward. In contrast with other responsible clinical research practices (such as data protection or trial registration itself), registration-publication linking is relatively straightforward and can be improved with action across stakeholders. <!--#Drawing from the study findings and additional context, -->Table \@ref(fig:tbl-stakeholders) outlines recommendations for stakeholders across clinical research.

```{r tbl-stakeholders}
# https://www.anthonyschmidt.co/post/2020-06-03-making-apa-tables-with-gt/
# results = "asis"
# tbl_stakeholders %>%
#   as_latex()
knitr::include_graphics(here::here("docs", "figures", "tbl-stakeholders.png"))
# tbl_stakeholders
```

@@

We would like to thank Kerstin Rubarth from the Charité Institut für Biometrie und Klinische Epidemiologie for her consultation on the statistical analysis, and Nicholas Devito for earlier versions of the trial registration number regular expressions.

@@

```{r render_appendix, include=FALSE}
render_appendix("appendix.Rmd")
```
